---
title: "tracks summary"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
    df_print: paged
---

## loading tracks, tracks_index, and index (cell information)
```{r, warning=FALSE, cache=TRUE}
library(tidyverse)
library(plotly) 
# plotly is only needed if looking at the interactive plots of the exponential 
# fit results, otherwise it can be commented out

# tracks path
all_tracks_path <- "../../data/tracks/all_tracks.rds"

all_tracks <- read_rds(all_tracks_path)

# spots_pre_during summary
spots_pre_during_step <- read_rds("../../data/tracks/pre_during_step_properties.rds")

# tracks index
tracks_index <- read_csv("../../data/index/tracks_index.csv")

# tracks <-
#   tracks %>%
#   left_join(tracks_index, by = c("base_filename", "episode", "track_id"))

# load master index for analysis
index <- 
  read_delim("../../data/index/high_speed_index.csv", delim = ",") %>%
  filter(
    protocol %in% 
      c("step50ms", "step50mspos", "step50msvid", "step50msvidavg4", "step50msposvidavg4", "step5msvidavg4"),
    track_mate_tracks == TRUE
  ) %>%
  select(-include, -best, -contains("notes_"), -contains("crop_"), -contains("gauss2d_"))

```

## expexponential fitting code
```{r} 
## exponential fitting code ==============================================

doubleExp = function(params, x) {
  # formula: y~ A0*exp(-x/tau0)) + A1*exp(-x/tau1)) + C0
  A0 = as.numeric(params["A0"])
  tau0 = as.numeric(params["tau0"])
  A1 = as.numeric(params["A1"])
  tau1 = as.numeric(params["tau1"])
  C0 = as.numeric(params["C0"])
  y = ((A0*exp(x/-tau0)) + (A1*exp(x/-tau1)) + C0)
  return(y)
}

singleExp = function(params, x) {
  # formula: y ~ A0*exp(-x/tau0)) + C0
  A0 = as.numeric(params["A0"])
  tau0 = as.numeric(params["tau0"])
  C0 = as.numeric(params["C0"])
  y = ((A0*exp(x/-tau0)) + (A1*exp(x/-tau1)) + C0)
  return(y)
}

doubleExpFit = function(x, y, startList, lower, upper, weightList){
  data = data.frame(x,y)
  fitModel <- 
    minpack.lm::nlsLM(y~((A0*exp(x/-tau0)) + (A1*exp(x/-tau1)) + C0),
                      data = data,
                      start = startList,
                      lower = lower,
                      upper = upper,
                      weights = weightList,
                      control = list(maxiter = 500,
                                     minFactor = 1/500))
  return(fitModel)
}


singleExpFit = function(x, y, startList, lower, upper, weightList){
  data = data.frame(x,y)
  fitModel <- 
    minpack.lm::nlsLM(y ~ (A1 * exp(x/-tau1)) + C0,
                      data = data,
                      start = startList,
                      lower = lower,
                      upper = upper,
                      weights = weightList,
                      control = list(maxiter = 500,
                                     minFactor = 1/500))
  return(fitModel)
}

tryDoubleExpFit <- function(x, y, startList, lower, upper, weightList) {
  out <- tryCatch(
    {
      doubleExpFit(x, y, startList, lower, upper, weightList)
      # The return value of `doubleExpFit()` is the actual value 
      # that will be returned in case there is no condition 
      # (e.g. warning or error).   
    },
    error = function(cond) {
      #print("in error part of try catch")
      #message("Cannot be fit")
      #message(cond)
      # Choose a return value in case of error
      return(NA_real_)
    },
    warning = function(cond) {
      #print("in warning part of try catch")
      #message("Warning")
      #message(cond)
      # Choose a return value in case of warning
      return(NA_real_)
    },
    finally = {
      # message("Some other message at the end")
    }
  )    
  return(out)
}

trySingleExpFit <- function(x, y, startList, lower, upper, weightList) {
  out <- tryCatch(
    {
      singleExpFit(x, y, startList, lower, upper, weightList)
      # The return value of `doubleExpFit()` is the actual value 
      # that will be returned in case there is no condition 
      # (e.g. warning or error).   
    },
    error = function(cond) {
      #print("in error part of try catch")
      #message("Cannot be fit")
      #message(cond)
      # Choose a return value in case of error
      return(NA_real_)
    },
    warning = function(cond) {
      #print("in warning part of try catch")
      #message("Warning")
      #message(cond)
      # Choose a return value in case of warning
      return(NA_real_)
    },
    finally = {
      # message("Some other message at the end")
    }
  )    
  return(out)
}
```

# Calculating peak displacements
```{r}
## Calculating summary data 
#===========================================================================

# Isochronal method - Probes
probes_peak <-
  all_tracks %>%
  left_join(
    index %>% select(base_filename, t_step_start), 
    by = c("base_filename")
  ) %>%
  dplyr::filter(
    track_id == 999,
    time > t_step_start + 0.15,
    time < t_step_start + 0.3
  ) %>%
  group_by(base_filename, episode) %>%
  summarise(
    probe_peak_nm_y = mean(position_nm_y_zeroed, na.rm = TRUE),
    probe_peak_nm_x = mean(position_nm_x_zeroed, na.rm = TRUE)
  ) %>%
  ungroup()

# Isochronal method - Stereocilia
# includes peak of all individual stereocilia (peak_nm_), 
# and the stereocilia with the largest displacement in the bundle (cilia_peak_nm_)
tracks_peak <-
  all_tracks %>%
  left_join(
    index %>% select(base_filename, t_step_start), 
    by = c("base_filename")
  ) %>%
  group_by(base_filename, episode, track_id) %>%  
  dplyr::filter(
    track_id != 999,
    low_noise == TRUE,
    include == TRUE,
    time > t_step_start + 0.15,
    time < t_step_start + 0.3
  ) %>%
  summarise(
    peak_nm_y = mean(position_nm_y_zeroed, na.rm = TRUE),
    peak_nm_x = mean(position_nm_x_zeroed, na.rm = TRUE),
    peak_nm_y_sd = sd(position_nm_y_zeroed, na.rm = TRUE),
    peak_nm_y_n = n(),
    peak_nm_y_se = peak_nm_y_sd / sqrt(peak_nm_y_n)
  ) %>%
  mutate(
    cilia_peak_nm_y = max(peak_nm_y, na.rm = TRUE),
    cilia_peak_nm_x = max(peak_nm_x, na.rm = TRUE)
  ) %>%
  ungroup()

# Maximum method, captures "peakiness" 
tracks_peak_max <-
  all_tracks %>%
  left_join(
    index %>% select(base_filename, t_step_start), 
    by = c("base_filename")
  ) %>%
  group_by(base_filename, episode, track_id) %>%
  dplyr::filter(
    track_id != 999,
    include == TRUE,
    low_noise == TRUE,
    time > t_step_start,
    time < t_step_start + 0.3
  ) %>% 
  summarise(
    peak_nm_y_max = max(position_nm_y_zeroed, na.rm = TRUE),
    time_peak_max = time[which.max(position_nm_y_zeroed)],
    peak_nm_x_max = max(position_nm_x_zeroed, na.rm = TRUE),
    time_peak_max_x = time[which.max(position_nm_x_zeroed)],
    peak_nm_y_max_sd = sd(position_nm_y_zeroed, na.rm = TRUE),
    peak_nm_y_max_n = n(),
    peak_nm_y_max_se = peak_nm_y_max_sd / sqrt(peak_nm_y_max_n)
  ) %>%
  mutate(
    cilia_peak_nm_y_max = max(peak_nm_y_max, na.rm = TRUE),
    cilia_peak_nm_x_max = max(peak_nm_x_max, na.rm = TRUE)
  ) %>%
  ungroup()

tracks_steady <-
  all_tracks %>%
  left_join(
    index %>% select(base_filename, t_step_end), 
    by = c("base_filename")
  ) %>%
  group_by(base_filename, episode, track_id) %>%
  dplyr::filter(
    track_id != 999,
    include == TRUE, 
    low_noise == TRUE,
    time > t_step_end - 1, 
    time < t_step_end
  ) %>%
  summarise(
    steady_nm_y = median(position_nm_y_zeroed, na.rm = TRUE),
    #steady_nm_y_mean = mean(position_nm_y_zeroed, na.rm = TRUE),
    #delta = median(position_nm_y_zeroed, na.rm = TRUE) - mean(position_nm_y_zeroed, na.rm = TRUE),
    steady_nm_x = median(position_nm_x_zeroed, na.rm = TRUE),
    steady_nm_y_n = n(),
    steady_nm_y_sd = sd(position_nm_y_zeroed, na.rm = TRUE),
    steady_nm_y_se = steady_nm_y_sd / sqrt(steady_nm_y_n)
  ) %>%
  ungroup() 

tracks_off_peak <-
  all_tracks %>%
  left_join(
    index %>% select(base_filename, t_step_end), 
    by = c("base_filename")
  ) %>%
  group_by(base_filename, episode, track_id) %>%  
  dplyr::filter(
    track_id != 999,
    include == TRUE,
    low_noise == TRUE,
    time > t_step_end,
    time < t_step_end + 0.4
  ) %>%
  summarise(
    off_peak_nm_y = min(position_nm_y_zeroed, na.rm = TRUE),
    time_off_peak = time[which.min(position_nm_y_zeroed)],
    off_peak_nm_x = min(position_nm_x_zeroed, na.rm = TRUE)
  ) %>%
  ungroup()  

tracks_off_steady <-
  all_tracks %>%
  left_join(
    index %>% select(base_filename, t_step_end), 
    by = c("base_filename")
  ) %>%
  group_by(base_filename, episode, track_id) %>%  
  dplyr::filter(
    track_id != 999,
    include == TRUE,
    low_noise == TRUE,
    time > t_step_end + 1,
    time < t_step_end + 5
  ) %>%
  summarise(
    off_steady_nm_y = mean(position_nm_y_zeroed, na.rm = TRUE),
    off_steady_nm_x = mean(position_nm_x_zeroed, na.rm = TRUE),
    off_steady_sd_y = sd(position_nm_y_zeroed, na.rm = TRUE),
    off_steady_sd_x = sd(position_nm_x_zeroed, na.rm = TRUE)
  ) %>%
  ungroup()

tracks_rise_inflection <-
  all_tracks %>%
  left_join(
    index %>% select(base_filename, t_step_start), 
    by = c("base_filename")
  ) %>%
  filter(
    track_id != 999,
    include == TRUE,
    low_noise == TRUE,
    time < t_step_start
  ) %>%
  group_by(base_filename, episode, track_id) %>%
  summarise(sd_d1 = sd(position_nm_y_zeroed_d1, na.rm = TRUE)) %>% 
  ungroup() %>%
  # ^ find the standard deviation of the first derivative of the baseline motion
  left_join(all_tracks, ., by = c("base_filename", "episode", "track_id")) %>%
  # ^ add all time points back 
  # ^ (this seems computationally intensive, maybe there's a way to add fewer points back?)
  left_join(
    tracks_peak %>% select(base_filename, episode, track_id, peak_nm_y), 
    by = c("base_filename", "episode", "track_id")
  ) %>%
  # ^ add the peak displacement for subsequent filtering
  left_join(
    index %>% select(base_filename, t_step_start), 
    by = c("base_filename")
  ) %>%
  # ^ add t_step_start for subsequent filtering
  group_by(base_filename, episode, track_id) %>%
  # ^ separate stereocilia tracks
  filter(
    peak_nm_y > 75,
    time > t_step_start, 
    time < t_step_start + .3,
    position_nm_y_zeroed_d1 > sd_d1
  ) %>%
  # ^ select only tracks where:
  # the peak is > 75nm, the time is during the onset of motion, and the first 
  # derivative is greater than the sd of the baseline first derivative
  top_n(1, position_nm_y_zeroed_d1) %>%
  # ^ take the row (point in time) with the steepest slope (highest first derivative of motion)
  ungroup() %>%
  group_by(base_filename, episode) %>%
  mutate(rise_delta_t_inflection = time - min(time)) %>%
  # ^ of all the stereocilia motion trajectory peak first derivatives in that episode, what is the difference in timing?
  rename(
    rise_t_inflection = time,
    rise_peak_slope = position_nm_y_zeroed_d1
  ) %>%
  ungroup() %>%
  select(
    base_filename, 
    episode, 
    track_id, 
    rise_t_inflection, 
    rise_delta_t_inflection, 
    rise_peak_slope
  ) %>%
  ungroup()
```

## New way of calculating timing:
Time of each stereocilium to some given distance. 
Saves to a separate onset timing data frame to access in figure_timing_v2.Rmd
```{r}
tracks_timing <-
  all_tracks %>%
  left_join(
    index %>% select(base_filename, t_step_start),
    by = c("base_filename")
  ) %>%
  filter(
    track_id != 999,
    include == TRUE,
    low_noise == TRUE,
    time < t_step_start + 1,
    time >= t_step_start
  )

tracks_timing <-
  tracks_timing %>%
  group_by(base_filename, episode, track_id) %>%
  summarise(sd_baseline = sd(position_nm_y_zeroed, na.rm = TRUE)) %>% 
  ungroup() %>%
  # ^ find the standard deviation of the baseline motion
  left_join(tracks_timing, ., by = c("base_filename", "episode", "track_id")) %>%
  left_join(
    tracks_peak %>% select(base_filename, episode, track_id, peak_nm_y), 
    by = c("base_filename", "episode", "track_id")
  ) %>%  
  # ^ add the peak displacement for subsequent filtering
  # group_by(base_filename, episode, track_id) %>%
  # # ^ separate stereocilia tracks
  filter(
    peak_nm_y > 75,
    time > t_step_start, 
    time < t_step_start + .5,
    peak_nm_y > sd_baseline
  )
  # ^ select only tracks where:
  # the peak is > 75nm, the time is during the onset of motion, and the peak 
  # is greater than the sd of the baseline motion

tracks_onset_interpolated <-
  tracks_timing %>%
  group_by(base_filename, episode, track_id) %>%
  summarise(
    interpolated = 
      approx(time, position_nm_y_zeroed, xout = seq(min(time), max(time), 0.001)) %>% 
      as.data.frame() %>% 
      list()
    ) %>%
  unnest() %>%
  rename(time = x, position_nm_y_zeroed = y) %>%
  ungroup()
  # ^ for each track, linearly interpolate displacements at 0.001 ms apart

tracks_timing_to_distance <-
  tracks_onset_interpolated %>%
  group_by(base_filename, episode, track_id) %>%
  filter(position_nm_y_zeroed <= 50) %>%
  top_n(1, time) %>%
  # ^ for each track, find the time at which the trajectory reaches 50 nm
  ungroup() %>%
  rename(t_to_50nm = time) 

tracks_timing_to_distance2 <-
  tracks_onset_interpolated %>%
  group_by(base_filename, episode, track_id) %>%
  filter(position_nm_y_zeroed <= 20) %>%
  top_n(1, time) %>%
  # ^ for each track, find the time at which the trajectory reaches 20 nm
  ungroup() %>%
  rename(t_to_20nm = time) 

tracks_timing_25_75 <-
  tracks_onset_interpolated %>%
  left_join(
    tracks_peak %>% select(base_filename, episode, track_id, peak_nm_y), 
    by = c("base_filename", "episode", "track_id")
  ) %>% 
  group_by(base_filename, episode, track_id) %>%
  # mutate(
  #   position_nm_y_zeroed_10p = peak_nm_y * 0.25, 
  #   position_nm_y_zeroed_90p = peak_nm_y * 0.75
  # ) %>%
  filter(
    position_nm_y_zeroed >= peak_nm_y * 0.25,
    position_nm_y_zeroed <= peak_nm_y * 0.75
  ) %>%
  slice(c(1,n())) %>%
  mutate(names = c("t_25", "t_75")) %>%
  ungroup() %>%
  # ^ for each track, find the time at which the trajectory reaches 10% and 90%
  #pivot_wider(names_from = names, values_from = time)
  select(-position_nm_y_zeroed) %>%
  spread(key = names, value = time) %>%
  mutate(t_rise_25_75 = t_75 - t_25)

tracks_onset_timings <-
  tracks_timing_to_distance %>%
  left_join(tracks_timing_to_distance2, by = c("base_filename", "episode", "track_id")) %>%
  left_join(tracks_timing_25_75, by = c("base_filename", "episode", "track_id"))

write_rds(
  tracks_onset_timings, 
  "/Volumes/GoogleDrive/My Drive/AlexTony/Projects/movementAnalysis/data/tracks_summary_onset.rds"
  )
write_csv(
  tracks_onset_timings, 
  "/Volumes/GoogleDrive/My Drive/AlexTony/Projects/movementAnalysis/data/tracks_summary_onset.csv"
  )

## Code used to figure out how to get timing at a given displacement
# test <-
#   tracks_onset_interpolated %>%
#   filter(
#     base_filename == tracks_timing$base_filename[[1]] | base_filename == tracks_timing$base_filename[[200]]  | base_filename == tracks_timing$base_filename[[400]] ,
#     episode == tracks_timing$episode[[1]] |  episode == tracks_timing$episode[[200]] |  episode == tracks_timing$episode[[400]],
#     track_id == tracks_timing$track_id[[1]] | track_id == tracks_timing$track_id[[200]] | track_id == tracks_timing$track_id[[400]]
#   ) %>%
#   group_by(base_filename, episode, track_id) %>%
#   filter(position_nm_y_zeroed <= 50) %>%
#   arrange(time)
# 
# test %>%
#   ggplot(aes(x = time, y = position_nm_y_zeroed, group = interaction(base_filename, episode, track_id))) +
#   geom_point()
# 
# test %>%
#   top_n(1, time) %>%
#     ggplot(aes(x = time, y = position_nm_y_zeroed, group = interaction(base_filename, episode, track_id))) +
#   geom_point()

## Code used to figure out how to do interpolation and add to data frame
# test <-
#   tracks_timing %>% 
#   filter(
#     base_filename == tracks_timing$base_filename[[1]] | base_filename == tracks_timing$base_filename[[200]] ,
#     episode == tracks_timing$episode[[1]] |  episode == tracks_timing$episode[[200]],
#     track_id == tracks_timing$track_id[[1]] | track_id == tracks_timing$track_id[[200]]
#   )
# 
# test_interp <- approx(test$time, test$position_nm_y_zeroed, xout = seq(min(test$time), max(test$time), 0.001))
# 
# test %>%
#   ggplot(aes(x = time, y = position_nm_y_zeroed, color = track_id)) +
#   geom_point()
# 
# test_interp %>%
#   as.data.frame() %>% nrow()
#   ggplot(aes(x = x, y = y)) +
#   geom_point()
#   
# test %>%
#   group_by(base_filename, episode, track_id) %>%
#   summarise(
#     interpolated = 
#       approx(time, position_nm_y_zeroed, xout = seq(min(time), max(time), 0.001)) %>% 
#       as.data.frame() %>% 
#       list()
#     ) %>%
#   unnest() %>%
#   ggplot(aes(x = x, y = y, group = interaction(base_filename, episode, track_id), color = as.factor(track_id))) +
#   geom_point()
```

## Combine tracks summaries & calculate mag reverse, percent reverse etc
Reverse motion is calculated if the peak displacement - steady state is greater than 2 x SD of the pre-step SD of movement AND peak motion is greater than 2 x SD pre step SD of movement (If you want to relax these requirements to look for the "peaky" reverse motions. Go with the max peak value and only 1 SD greater than noise)
```{r}
tracks_summary <-
  tracks_peak %>%
  left_join(tracks_peak_max, by = c("base_filename", "episode", "track_id")) %>%
  left_join(tracks_steady, by = c("base_filename", "episode", "track_id")) %>%
  left_join(tracks_off_peak, by = c("base_filename", "episode", "track_id")) %>%
  left_join(tracks_off_steady, by = c("base_filename", "episode", "track_id")) %>%
  left_join(probes_peak, by = c("base_filename", "episode")) %>%
  left_join(
    spots_pre_during_step %>% 
      select(base_filename, episode, track_id, prestep_sd_y, step_sd_y, prestep_sd_x, step_sd_x), 
    by = c("base_filename", "episode", "track_id")
    ) %>%
  mutate(
    mag_reverse = 
      case_when(
        peak_nm_y_max < (2 * prestep_sd_y) |
          (peak_nm_y_max - steady_nm_y) < (2 * step_sd_y) | 
          steady_nm_y < 0 - (3 * prestep_sd_y) ~ 
          0, 
        peak_nm_y_max >= (2 * prestep_sd_y) & 
          (peak_nm_y_max - steady_nm_y) >=  (2 * step_sd_y) & 
          steady_nm_y >= 0 - (3 * prestep_sd_y) ~ 
          peak_nm_y_max - steady_nm_y,
        TRUE ~ 
          NA_real_
      ),
    percent_reverse = mag_reverse / peak_nm_y_max,
    mag_reverse_x = 
      case_when(
        peak_nm_x_max < (2 * prestep_sd_x) | 
          (peak_nm_x_max - steady_nm_x) < (2 * step_sd_x) | 
          steady_nm_x < 0 - (3 * prestep_sd_x) ~ 
          0, 
        peak_nm_x_max >= (2 * prestep_sd_x) & 
          (peak_nm_x_max - steady_nm_x) >= (2 * step_sd_x) & 
          steady_nm_x >= 0 - (3 * prestep_sd_x) ~ 
          peak_nm_x_max - steady_nm_x,
        TRUE ~ 
          NA_real_
      ),
    percent_reverse_x = mag_reverse_x / peak_nm_x_max
  ) %>%
  mutate(
    off_mag = steady_nm_y - off_peak_nm_y,
    off_notch = 
      case_when(
        off_peak_nm_y < off_steady_nm_y & 
          abs(off_peak_nm_y) > (2 * off_steady_sd_y) & 
          off_steady_nm_y < (3 * prestep_sd_y) ~ 
          abs(off_peak_nm_y - off_steady_nm_y),
        off_peak_nm_y >= off_steady_nm_y | 
          abs(off_peak_nm_y) <= (2 * off_steady_sd_y) | 
          off_steady_nm_y >= (3 * prestep_sd_y) | 
          (off_peak_nm_y > 0) ~ 
          0,
        TRUE ~ 
          NA_real_
      )
  ) %>%
  left_join(tracks_rise_inflection, by = c("base_filename", "episode", "track_id")) %>%
  left_join(
    tracks_index,
    by = c("base_filename", "episode", "track_id")
  ) %>%
  left_join(
    index,
    by = c("base_filename")
  ) %>%
  mutate(
    holding = if_else(str_detect(protocol, "pos"), 80, -80),
    contact =
      case_when(
        direct_contact == TRUE ~ "direct",
        indirect_contact == TRUE ~ "indirect",
        TRUE ~ "none"
      ),
    side = if_else(position_id > 0, "right", "left"),
    side = if_else(position_id == 0, "center", side)
    ) %>%
  filter(include == TRUE, is_probe == FALSE) %>%
  select(
    date, age, sex, location, probe,
    cell_id, folder_name,
    base_filename, episode, 
    focal_row, cilium_row, 
    track_id, position_id,
    everything()
  ) %>%
  select(
    -include, -track_mate_tracks, -track_index, -I_peak_pA, -leak_pA, -cm, -rs, 
    -step_size_mV, -um_mV, -direct_contact, -indirect_contact, -bundle_row, -is_probe
  )
```


# Add fits
Fitting first 25 ms
```{r}
tracks_fits_1 <- # fit 25 ms
  all_tracks %>%
  inner_join(
    tracks_summary %>% 
      select(base_filename, episode, track_id, mag_reverse, time_peak_max, t_step_start, t_step_end) %>%
      filter(mag_reverse > 0),
    by = c("base_filename", "episode", "track_id")
  ) %>%
  #single_track %>%
  mutate(
    #t_fit_start = t_step_start + .2, # for isochronal peak, add amount of time to peak
    t_fit_start = time_peak_max, # use this if max was taken at peak
    t_fit_end = map2_dbl(t_step_end, t_step_start, ~ min(.x, .y + 25, na.rm = TRUE))
  ) %>%
  filter(
    time > t_fit_start, 
    time < t_fit_end
  ) %>%
  group_by(base_filename, episode, track_id, t_fit_start, t_fit_end) %>%  
  summarise(
    fit_dbl = 
      tryDoubleExpFit(
        x = time - first(time), 
        y = position_nm_y_zeroed,
        startList = 
          list(
            A0 = (first(position_nm_y_zeroed) - last(position_nm_y_zeroed)) / 2,
            #tau0 = (last(time) - first(time)) / 5,
            tau0 = 5,
            A1 = (first(position_nm_y_zeroed) - last(position_nm_y_zeroed)) / 2,
            #tau1 = (last(time) - first(time)) / 20,
            tau1 = .5,
            C0 = last(position_nm_y_zeroed)
          ),
        lower = c(0, 0, 0, 0, 0), # limits c(A0, tau0, A1, tau1, C0)
        upper = 
          c(
            1000, 
            1000, #(last(time) - first(time)) / 3, 
            1000, 
            1000, #(last(time) - first(time)) / 3, 
            2000
          ), # limits c(A0, tau0, A1, tau1, C0)
        weightList = 
          rep(1, length(position_nm_y_zeroed))
          #exp(-(seq(1:length(position_nm_y_zeroed))) / (length(position_nm_y_zeroed) / 15))
      ) %>%
      list(),
    fit_sgl = 
      trySingleExpFit(
        x = time - first(time), 
        y = position_nm_y_zeroed,
        startList = 
          list(
            A1 = (first(position_nm_y_zeroed) - last(position_nm_y_zeroed)) / 2,
            #tau1 = (last(time) - first(time)) / 5,
            tau1 = .5,
            C0 = last(position_nm_y_zeroed)
          ),
        lower = c(0, 0, 0), # limits c(A1, tau1, C0)
        upper = 
          c(
            1000, 
            1000, #(last(time) - first(time)) / 3, 
            2000
          ), # limits c(A0, tau0, A1, tau1, C0)
        weightList = 
          rep(1, length(position_nm_y_zeroed))
          #exp(-(seq(1:length(position_nm_y_zeroed))) / (length(position_nm_y_zeroed) / 1))
      ) %>%
      list(),
    predict_time = list(time)
  ) %>%
  ungroup() %>%
  mutate(    
    predict_dbl = map(fit_dbl, ~ if(!is.na(.[1])) {predict(.)} else {NA_real_}),
    predict_sgl = map(fit_sgl, ~ if(!is.na(.[1])) {predict(.)} else {NA_real_})
  )

tracks_fits_2 <- # fit 5ms
  all_tracks %>%
  inner_join(
    tracks_summary %>% 
      select(base_filename, episode, track_id, mag_reverse, time_peak_max, t_step_start, t_step_end) %>%
      filter(mag_reverse > 0),
    by = c("base_filename", "episode", "track_id")
  ) %>%
  #single_track %>%
  mutate(
    #t_fit_start = t_step_start + .2, # for isochronal peak, add amount of time to peak
    t_fit_start = time_peak_max, # use this if max was taken at peak
    t_fit_end = map2_dbl(t_step_end, t_step_start, ~ min(.x, .y + 2, na.rm = TRUE))
  ) %>%
  filter(    
    time > t_fit_start,
    time < t_fit_end
  ) %>%
  group_by(base_filename, episode, track_id, t_fit_start, t_fit_end) %>%  
  summarise(
    fit_sgl_short = 
      trySingleExpFit(
        x = time - first(time), 
        y = position_nm_y_zeroed,
        startList = 
          list(
            A1 = (first(position_nm_y_zeroed) - last(position_nm_y_zeroed)) / 2,
            #tau1 = (last(time) - first(time)) / 20,
            tau1 = .5,
            C0 = last(position_nm_y_zeroed)
          ),
        lower = c(0, 0, 0), # limits c(A1, tau1, C0)
        upper = 
          c(
            1000, 
            1000, #(last(time) - first(time)) / 3, 
            2000
          ), # limits c(A0, tau0, A1, tau1, C0)
        weightList = 
          rep(1, length(position_nm_y_zeroed))
          #exp(-(seq(1:length(position_nm_y_zeroed))) / (length(position_nm_y_zeroed) / 1))
      ) %>%
      list(),
    predict_time_short = list(time)
  ) %>%
  ungroup() %>%
  mutate(
    predict_sgl_short = map(fit_sgl_short, ~ if(!is.na(.[1])) {predict(.)} else {NA_real_})
  ) %>%
  select(base_filename, episode, track_id, fit_sgl_short, predict_time_short, predict_sgl_short)

tracks_fits_off <-
  all_tracks %>%
  inner_join(
    tracks_summary %>% 
      select(base_filename, episode, track_id, off_notch, time_off_peak) %>%
      filter(off_notch > 0),
    by = c("base_filename", "episode", "track_id")
  ) %>%
  #single_track %>%
  mutate(
    t_fit_start = time_off_peak, # use this if max was taken at peak
    t_fit_end = time_off_peak + 3 # previous to 6/19, did 5 ms
  ) %>%
  filter(
    time > t_fit_start, 
    time < t_fit_end
  ) %>%
  group_by(base_filename, episode, track_id, t_fit_start, t_fit_end, off_notch, time_off_peak) %>%  
  summarise(
    fit_sgl_off = 
      trySingleExpFit(
        x = time - first(time), 
        y = position_nm_y_zeroed,
        startList = 
          list(
            A1 = (first(position_nm_y_zeroed) - last(position_nm_y_zeroed)) / 2,
            tau1 = .5,
            C0 = last(position_nm_y_zeroed)
          ),
        lower = c(-5000, 0, -5000), # limits c(A1, tau1, C0)
        upper = c(0, 10, 2000), # limits c(A1, tau1, C0)
        weightList = 
          rep(1, length(position_nm_y_zeroed))
      ) %>%
      list(),
    predict_time_off = list(time)
  ) %>%
  ungroup() %>%
  mutate(
    predict_sgl_off = map(fit_sgl_off, ~ if(!is.na(.[1])) {predict(.)} else {NA_real_})
  ) %>%
  select(base_filename, episode, track_id, time_off_peak, off_notch, fit_sgl_off, predict_time_off, predict_sgl_off)

# # weight list
# x = seq(1:100)
# y = exp(-(seq(1:100)) / (100 / 1))
# plot(x, y)
```

```{r}
#tracks_fits_2$fit_sgl[[1]] %>% summary()

#tau1 is fast
#tau0 is slow

tracks_fits <-
  tracks_fits_1 %>%
  left_join(tracks_fits_2, by = c("base_filename", "episode", "track_id")) %>%
  left_join(tracks_fits_off, by = c("base_filename", "episode", "track_id")) %>%
  mutate(
    tau1_sgl = map_dbl(fit_sgl, ~ if(!is.na(.[1])) {coef(.)["tau1"]} else {NA_real_}),
    tau1_sgl_short = map_dbl(fit_sgl_short, ~ if(!is.na(.[1])) {coef(.)["tau1"]} else {NA_real_}),
    tau1_sgl_off = map_dbl(fit_sgl_off, ~ if(is.list(.[1])) {coef(.)["tau1"]} else {NA_real_}),
    
    tau1_dbl = map_dbl(fit_dbl, ~ if(!is.na(.[1])) {coef(.)["tau1"]} else {NA_real_}),  
    tau0_dbl = map_dbl(fit_dbl, ~ if(!is.na(.[1])) {coef(.)["tau0"]} else {NA_real_}),
    
    # record whether tau1 goes with fast for slow (for setting As later)
    tau1_dbl_fast = if_else(tau1_dbl < tau0_dbl, TRUE, FALSE), 
    
    # Set fast time constant to be fast, and slow to be slow
    tau1_dbl = pmin(tau1_dbl, tau0_dbl), 
    tau0_dbl = pmax(tau1_dbl, tau0_dbl),
    
    # getting A0 and A1 parameters
    A1_sgl = map_dbl(fit_sgl, ~ if(!is.na(.[1])) {coef(.)["A1"]} else {NA_real_}),
    A1_sgl_off = map_dbl(fit_sgl_off, ~ if(is.list(.[1])) {coef(.)["A1"]} else {NA_real_}),
    A1_dbl = map_dbl(fit_dbl, ~ if(!is.na(.[1])) {coef(.)["A1"]} else {NA_real_}),
    A0_dbl = map_dbl(fit_dbl, ~ if(!is.na(.[1])) {coef(.)["A0"]} else {NA_real_}),
    
    #make sure double fit As match the correct time constant
    A1_dbl = if_else(tau1_dbl_fast == TRUE, A1_dbl, A0_dbl),
    A0_dbl = if_else(tau1_dbl_fast == TRUE, A0_dbl, A1_dbl)
    
    #resid_sd_sgl = map_dbl(fit_sgl, ~ if(!is.na(.[1])) {sigma(.)} else {NA_real_}),
    #resid_sd_dbl = map_dbl(fit_dbl, ~ if(!is.na(.[1])) {sigma(.)} else {NA_real_})
  ) %>%
  mutate(
    best_fit = 
      case_when(
        
        # if its a single AND NOT a double
        # AND single fit tau 1 is fast enough
        # ---OR---
        # if fits a single AND fits a double
        # AND single fit tau1 is fast enough
        # AND double fit has too slow a tau0 OR (EXCLUDING TOO SLOW REASON FOR NOW)
        #   too fast a tau1 OR (EXCLUDING TOO FAST REASON FOR NOW)
        #   the step is very short OR (EXCLUDING TOO SHORT REASON FOR NOW)
        #   both tau0 and tau1 of the double fit are less than 1ms different
        # THEN the best fit is single
        
        (!(fit_sgl %>% is.na()) & (fit_dbl %>% is.na()) & 
           tau1_sgl < ((t_fit_end - t_fit_start) / 3)) |
        # ---OR---
        (!(fit_sgl %>% is.na()) & !(fit_dbl %>% is.na()) & 
          tau1_sgl < ((t_fit_end - t_fit_start) / 3) &
          #(tau0_dbl > ((t_fit_end - t_fit_start) / 3) | 
          (tau1_dbl < 0.001 | 
             #t_fit_end - t_fit_start < 15 |
             abs(tau0_dbl - tau1_dbl) < 1))
          ~ "single", 
        
        # if its a double AND NOT a single
        # AND double fit has fast enough tau0 AND (EXCLUDING TOO FAST REASON FOR NOW)
        #   slow enough a tau1 AND (EXCLUDING TOO SLOW REASON FOR NOW)
        #   the step is long enough AND
        #   both tau0 and tau1 of the double fit are more than than 2 ms different
        # ---OR---
        # if fits a single AND fits a double
        # AND double fit has fast enough tau0 AND  (EXCLUDING TOO FAST REASON FOR NOW)
        #   slow enough a tau1 AND (EXCLUDING TOO SLOW REASON FOR NOW)
        #   the step is long enough AND (EXCLUDING TOO SHORT REASON FOR NOW)
        #   both tau0 and tau1 of the double fit are more than than 2 ms different
        
        (!(fit_dbl %>% is.na()) & (fit_sgl %>% is.na()) &
          tau0_dbl < t_fit_end - t_fit_start &
           #(tau0_dbl < ((t_fit_end - t_fit_start) / 3) & 
           #(tau1_dbl > 0.001 & 
           # (t_fit_end - t_fit_start > 15 &
              (abs(tau0_dbl - tau1_dbl) > 1)) |
        # ---OR---
        (!(fit_sgl %>% is.na()) & !(fit_dbl %>% is.na()) & 
           tau0_dbl < t_fit_end - t_fit_start &
          #(tau0_dbl < ((t_fit_end - t_fit_start) / 3) & 
          #(tau1_dbl > 0.001 & 
          #(t_fit_end - t_fit_start > 15) &
              (abs(tau0_dbl - tau1_dbl) > 1))
          ~ "double", 
        TRUE ~ NA_character_
      ),
    best_fit = # second round of best fit determination
      case_when(
        best_fit == "double" & 
          tau0_dbl > 50 &
          tau1_sgl_short < ((t_fit_end - t_fit_start) / 3) ~ "single_short",
        is.na(best_fit) &
          tau1_sgl_short < ((t_fit_end - t_fit_start) / 3) ~ "single_short",
        TRUE ~ best_fit
      ),
    tau_fast = 
      case_when(
      best_fit == "single" ~ tau1_sgl,
      best_fit == "double" ~ tau1_dbl,
      best_fit == "single_short" ~ tau1_sgl_short,
      TRUE ~ NA_real_
    ),
    tau_slow = 
      case_when(
        best_fit == "double" ~ tau0_dbl,
        TRUE ~ NA_real_
      ),
    tau_off = case_when(
      tau1_sgl_off < 10 ~ tau1_sgl_off,
      TRUE ~ NA_real_
    )
  )


```

# plot fits against data
```{r, eval = FALSE}
plot_raw_pred <- function(b, e, t, best){
  curr_fit <-         
    tracks_fits %>%
        filter(
          base_filename == b, 
          episode == e, 
          track_id == t
        )
  
  df_predict_1 <-
    tibble(
      time = curr_fit %>% select(predict_time) %>% unlist(),
      predict_1 = curr_fit %>% select(predict_sgl) %>% unlist()
    )
  
  df_predict_2 <-
    tibble(
      time = curr_fit %>% select(predict_time) %>% unlist(),
      predict_2 = curr_fit %>% select(predict_dbl) %>% unlist()
    )
  
  df_predict_3 <-
    tibble(
      time = curr_fit %>% select(predict_time_short) %>% unlist(),
      predict_3 = curr_fit %>% select(predict_sgl_short) %>% unlist()
    )
  
  df_raw <-
    all_tracks %>% 
    filter(
      base_filename == b, 
      episode == e, 
      track_id == t
    ) %>% 
    select(time, position_nm_y_zeroed)
  
  p <- ggplot() +
    geom_line(data = df_raw, aes(x = time, y = position_nm_y_zeroed)) +
    geom_line(data = df_predict_2, aes(x = time, y = predict_2), color = "red") +
    geom_line(data = df_predict_1, aes(x = time, y = predict_1), color = "yellow") +
    geom_line(data = df_predict_3, aes(x = time, y = predict_3), color = "green") +
    labs(title = str_c(b, " - ", t, " - ", paste(best)))#, subtitle = paste(best))
  ggplotly(p)
}

tracks_fits[sample(seq(1:nrow(tracks_fits)), 10),] %>%
  select(base_filename, episode, track_id, best_fit) %>%
  pmap(~ plot_raw_pred(b = ..1, e = ..2, t = ..3, best = ..4))  

#plot_raw_pred(b = "S8731024", e = 1, t = 8, b = "double")

#tracks_fits[2,]$fit_dbl[[1]] %>% summary()

tracks_fits %>%
  filter(base_filename == "S8731024", episode == 1, track_id == 29) %>%
  select(base_filename, episode, track_id, best_fit) %>%
  pmap(~ plot_raw_pred(b = ..1, e = ..2, t = ..3, best = ..4))  

```

# plot fits against data - off notch
```{r, eval = FALSE}
plot_raw_pred <- function(b, e, t){
  curr_fit <-         
    tracks_fits_off %>%
        filter(
          base_filename == b, 
          episode == e, 
          track_id == t
        )
  
  df_predict <-
    tibble(
      time = curr_fit %>% select(predict_time_off) %>% unlist(),
      predict = curr_fit %>% select(predict_sgl_off) %>% unlist()
    )
  
  df_raw <-
    all_tracks %>% 
    filter(
      base_filename == b, 
      episode == e, 
      track_id == t
    ) %>% 
    select(time, position_nm_y_zeroed)
  
  p <- ggplot() +
    geom_line(data = df_raw, aes(x = time, y = position_nm_y_zeroed)) +
    geom_line(data = df_predict, aes(x = time, y = predict), color = "red") +
    geom_vline(data = curr_fit, aes(xintercept = time_off_peak), color = "blue") +
    labs(title = str_c(b, " - ", t))#, subtitle = paste(best))
  ggplotly(p)
}

curr_df <-
  tracks_fits_off %>%
  filter(off_notch > 0)

curr_df[sample(seq(1:nrow(curr_df)), 10),] %>%
  select(base_filename, episode, track_id) %>%
  pmap(~ plot_raw_pred(b = ..1, e = ..2, t = ..3))  

# tracks_fits_off %>%
#   filter(base_filename == "S8731024", episode == 1, track_id == 29) %>%
#   select(base_filename, episode, track_id) %>%
#   pmap(~ plot_raw_pred(b = ..1, e = ..2, t = ..3))  

```

```{r}
tracks_summary_1 <-
  tracks_summary %>%
  left_join(
    tracks_fits %>% select(-off_notch, -time_off_peak), 
    by = c("base_filename", "episode", "track_id")
  ) %>%
  #select(-time) %>% 
  ungroup() %>%
  mutate(
    displacement_include = 
      case_when(
        cell_id == "20180731b2" & is.na(push_location) ~ "TRUE", # probe is touching stereocilia -1 and 0, both sides can be used
        cell_id == "20180731c4" & is.na(push_location) & position_id <= -1 ~ "TRUE", # both sides, -1 and 2
        cell_id == "20180731c4" & is.na(push_location) & position_id >= 2 ~ "TRUE",
        cell_id == "20180731c4" & push_location == "right" & position_id <= 5 ~ "TRUE",
        cell_id == "20180731c4" & push_location == "left" & position_id >= -4 ~ "TRUE",
        cell_id == "20180802b2" & is.na(push_location) & position_id <= 0 ~ "TRUE",
        cell_id == "20180802c2" & is.na(push_location) ~ "TRUE", # both sides, 0 and 1
        cell_id == "20180803a1" & is.na(push_location) & position_id >= 1 ~ "TRUE",
        cell_id == "20180803a1" & push_location == "right" & position_id <= 6 ~ "TRUE",
        cell_id == "20180804b3" & is.na(push_location) & position_id <= -1 ~ "TRUE",
        cell_id == "20180807c3" & is.na(push_location) ~ "TRUE", # probe is really just touching center stereocilium, both sides can be used
        cell_id == "20180807c3" & push_location == "left" & position_id >= -6 ~ "TRUE",
        cell_id == "20180808a3" & is.na(push_location) & position_id <= -1 ~ "TRUE",
        cell_id == "20180809b2" & is.na(push_location) & position_id <= -1 ~ "TRUE",
        # cell_id == "20180809b3" ~ "FALSE", # probe-touched-stereocilia were not tracked. No accurate 0 position
        # cell_id == "20180810a1" ~ "FALSE", # overlapped stereocilia in middle, missing 0 position
        cell_id == "20180811a1" & is.na(push_location) & position_id <= -1 ~ "TRUE",
        cell_id == "20180811a2" & is.na(push_location) ~ "TRUE", # both sides, -1 and 0,
        # cell_id == "20180811b3" ~ "FALSE", # maybe left side, starting at -1 would be ok
        cell_id == "20180815b3" & is.na(push_location) ~ "TRUE", # both sides, around -1 and 1,
        cell_id == "20180815b3" & push_location == "left" & position_id >= -3 ~ "TRUE", # to the right of -3,
        cell_id == "20180815b3" & push_location == "right" & position_id <= 7 ~ "TRUE" # to the left of 7
      ),
    displacement_edge = 
      case_when(
        cell_id == "20180731b2" & is.na(push_location) & side == "left" ~ -1,
        cell_id == "20180731b2" & is.na(push_location) & side != "left" ~ 0,
        cell_id == "20180731c4" & is.na(push_location) & side == "left" ~ -1,
        cell_id == "20180731c4" & is.na(push_location) & side == "right" ~ 2,
        cell_id == "20180731c4" & push_location == "right" ~ 5,
        cell_id == "20180731c4" & push_location == "left" ~ -4,
        cell_id == "20180802b2" & is.na(push_location) ~ 0,
        cell_id == "20180802c2" & is.na(push_location) & side != "right" ~ 0, # both sides, 0 and 1
        cell_id == "20180802c2" & is.na(push_location) & side == "right" ~ 1, # both sides, 0 and 1
        cell_id == "20180803a1" & is.na(push_location) ~ 1,
        cell_id == "20180803a1" & push_location == "right" ~ 6,
        cell_id == "20180804b3" & is.na(push_location) ~ -1,
        cell_id == "20180807c3" & is.na(push_location) ~ 0, # 0 is last touched for both sides
        cell_id == "20180807c3" & push_location == "left" ~ -5, 
        cell_id == "20180808a3" & is.na(push_location) ~ -1,
        cell_id == "20180809b2" & is.na(push_location) ~ -1,
        cell_id == "20180811a1" & is.na(push_location) ~ -1,
        cell_id == "20180811a2" & is.na(push_location) & side == "left" ~ -1,
        cell_id == "20180811a2" & is.na(push_location) & side != "left" ~ 0,
        cell_id == "20180815b3" & is.na(push_location) & side == "left" ~ -1, 
        cell_id == "20180815b3" & is.na(push_location) & side != "left" ~ 1,
        cell_id == "20180815b3" & push_location == "left" ~ -3,
        cell_id == "20180815b3" & push_location == "right" ~ 7
      ),
    normalized_side = 
      case_when(
        cell_id == "20180731b2" & is.na(push_location) & side == "left" ~ "left",
        cell_id == "20180731b2" & is.na(push_location) & side != "left" ~ "right",
        cell_id == "20180731c4" & is.na(push_location) & side == "left" ~ "left",
        cell_id == "20180731c4" & is.na(push_location) & side == "right" ~ "right",
        cell_id == "20180731c4" & push_location == "right" ~ "left",
        cell_id == "20180731c4" & push_location == "left" ~ "right",
        cell_id == "20180802b2" & is.na(push_location) & side == "left" ~ "left",
        cell_id == "20180802b2" & is.na(push_location) & side != "left" ~ "right",
        cell_id == "20180802c2" & is.na(push_location) & side != "right" ~ "left", # both sides, 0 and 1
        cell_id == "20180802c2" & is.na(push_location) & side == "right" ~ "right", # both sides, 0 and 1
        cell_id == "20180803a1" & is.na(push_location) ~ "right",
        cell_id == "20180803a1" & push_location == "right" ~ "left",
        cell_id == "20180804b3" & is.na(push_location) ~ "left",
        cell_id == "20180807c3" & is.na(push_location) & side == "left" ~ "left", # 0 is last touched for both sides
        cell_id == "20180807c3" & is.na(push_location) & side != "left" ~ "right", # 0 is last touched for both sides
        cell_id == "20180807c3" & push_location == "left" ~ "right",
        cell_id == "20180808a3" & is.na(push_location) ~ "left",
        cell_id == "20180809b2" & is.na(push_location) ~ "left",
        cell_id == "20180811a1" & is.na(push_location) ~ "left",
        cell_id == "20180811a2" & is.na(push_location) & side == "left" ~ "left",
        cell_id == "20180811a2" & is.na(push_location) & side != "left" ~ "right",
        cell_id == "20180815b3" & is.na(push_location) & side == "left" ~ "left", 
        cell_id == "20180815b3" & is.na(push_location) & side != "left" ~ "right",
        cell_id == "20180815b3" & push_location == "left" ~ "right",
        cell_id == "20180515b3" & push_location == "right" ~ "left"
      )
  ) %>%
  group_by(base_filename, episode) %>%
  mutate(
    norm_position_id = 
           if_else(displacement_include == TRUE, position_id - displacement_edge, NA_real_)
  ) %>%
  ungroup()

write_rds(tracks_summary_1, "../../data/tracks_summary.rds")
```
